---
title: "Tooth Works"
author: "Paul Hickman"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 1
  html_document:
    keep_md: false
---

# Overview

Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 
- Load the ToothGrowth data and perform some basic exploratory data analyses 
- Provide a basic summary of the data.
- Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
- State your conclusions and the assumptions needed for your conclusions. 

Some criteria that you will be evaluated on
- Did you  perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
- Did the student perform some relevant confidence intervals and/or tests?
- Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 
- Did the student describe the assumptions needed for their conclusions?

# Start

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

A data frame with 60 observations on 3 variables.

[,1]	len	numeric	Tooth length
[,2]	supp	factor	Supplement type (VC or OJ).
[,3]	dose	numeric	Dose in milligrams/day

https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html



```{r}
data("ToothGrowth")

str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
table(ToothGrowth$supp,ToothGrowth$dose)

# RR relationships between predictors
#library(kernlab)
#data(spam)
##structureOfADataAnalysis2.pdf
#plot(log10(trainSpam$capitalAve + 1) ~ trainSpam$type)
#plot(log10(trainSpam[, 1:4] + 1))
#plot(hclust(dist(t(trainSpam[,1:57]))))
#plot(hclust(dist(t(log10(trainSpam[,1:55]))))

plot(log10(ToothGrowth$len))
plot(log10(ToothGrowth$dose))

boxplot(len ~ supp * dose, data=ToothGrowth, ylab="Tooth Length", main="Boxplots of Tooth Growth Data")

require(graphics)
coplot(len ~ dose | supp, data = ToothGrowth, panel = panel.smooth)

library(ggplot2)
ggplot(ToothGrowth, aes(x = supp, y = len)) + geom_boxplot()
ggplot(ToothGrowth, aes(x = supp, y = len, fill = supp)) + geom_boxplot()
ggplot(ToothGrowth, aes(x = supp, y = len, fill = supp)) + geom_boxplot() + facet_wrap(~ dose)
```


# Techo
```{r}
sessionInfo()
```

