---
title: "Investigate the effect of Vitamin C on Guinea Pig tooth growth"
author: "Paul Hickman"
date: '`r Sys.Date()`'
output:
  pdf_document:
    default
  html_document:
    keep_md: no
---

# Overview

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html

# Data Summary

Load the tooth growth data that will be analysed.

```{r}
library(datasets)
data("ToothGrowth")
```

The data consists of 60 observations on 3 variables.

* [,1]	len	numeric	Tooth length
* [,2]	supp	factor	Supplement type (VC or OJ).
* [,3]	dose	numeric	Dose in milligrams/day

```{r}
summary(ToothGrowth)
```

There are three dosage levels (0.5, 1 & 2 mg/d) for the two supplement types orange juice (OJ) and ascorbic acid (VC) used.

# Exploratory Analysis

The primary question to answer is which supplement encourages more growth and which dose level.

First, evaluate the supplement/dose on tooth length.

```{r}
library(ggplot2)
# make dose a factor
ToothGrowth$dose <- factor(ToothGrowth$dose)

# supplement by dose
ggplot(ToothGrowth, aes(x = supp, y = len, fill = supp)) + geom_boxplot() + 
    facet_wrap(~dose) + xlab("Supplement") + ylab("Tooth Length") + 
    ggtitle("Length by Supplement and Dose")
```

Second, evaluate the dose/supplement on tooth length.

```{r}
# dose by supplement
ggplot(ToothGrowth, aes(x = dose, y = len, fill = dose)) + geom_boxplot() + 
    facet_grid(~supp) + xlab("Tooth Length") + ylab("Supplement") + 
    ggtitle("Length by Dose and Supplement")
```

The exploratory plots indicate the higher the dose the greater the tooth length.  Also, orange juice is associated with greater tooth length, except for the 2mg/g ascorbic acid (VC) supplement.

# Tests

The null hypothesis for the tests is the supplement type and dose have no effect on tooth growth.  Details of the tests performed can be found in the appendix.

The tests indicate that higher doses, no matter the supplement, is associated with greater tooth length.  However, OJ rather than VC is associated with greater tooth length for 0.5 and 1 mg/d dose and is not sufficiently different at the 2 mg/d level.

**Assumptions**

* The variance between the groups is the same
* There are no other contributing factors to tooth growth
* The sample size is adequate

# Appendix

To compare the VC supplement against the three dosage levels we need to subset the data first.

```{r}
# by supplement
vc <- ToothGrowth[ToothGrowth$supp=="VC",]

vc_05_1 <- vc[vc$dose %in% c(0.5,1),]
vc_05_2 <- vc[vc$dose %in% c(0.5,2),]
vc_1_2 <- vc[vc$dose %in% c(1,2),]

t.test(len ~ dose, data = vc_05_1)
t.test(len ~ dose, data = vc_05_2)
t.test(len ~ dose, data = vc_1_2)
```

The null hypothesis that there is no difference can be rejected in each case.  The higher the VC dose the greater the tooth length.

To compare the OJ supplement against the three dosage levels we need to subset the data first.

```{r}
# by supplement
oj <- ToothGrowth[ToothGrowth$supp=="OJ",]

oj_05_1 <- oj[oj$dose %in% c(0.5,1),]
oj_05_2 <- oj[oj$dose %in% c(0.5,2),]
oj_1_2 <- oj[oj$dose %in% c(1,2),]

t.test(len ~ dose, data = oj_05_1)
t.test(len ~ dose, data = oj_05_2)
t.test(len ~ dose, data = oj_1_2)
```

The null hypothesis that there is no difference can be rejected in each case.  The higher the OJ dose the greater the tooth length.

To compare the dosage to supplement we need to subset the data first.

```{r}
# by dose
tg05 <- ToothGrowth[ToothGrowth$dose == 0.5,]
tg1 <- ToothGrowth[ToothGrowth$dose == 1,]
tg2 <- ToothGrowth[ToothGrowth$dose == 2,]

t.test(len ~ supp, data = tg05)
t.test(len ~ supp, data = tg1)
t.test(len ~ supp, data = tg2)
```

The null hypothesis that there is no difference between supplements can be rejected in each case, except for dose 2mg/d as the confidence interval includes zero.  For dose 0.5 and 1 OJ is associated with greater tooth length.
