---
title: "A comparison of Exponential Distribution and the Central Limit Theorem"
author: "Paul Hickman"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    default
  html_document:
    keep_md: false
---

## Overview

Investigate the random exponential distribution and the Central Limit Theorem.  The Central Limit Theorem states that the distribution of averages of IID variables (properly normalised) becomes that of a standard normal as the sample size increases.  This investigation should demonstrate the average random exponential sample to be normally distributed.

## Simulations

The exponential distribution is simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. For this investigation lambda is 0.2 for all of the simulations.  The simulation will create a distribution of averages of 40 exponentials 1000 times.

```{r}
# number of simulations, lambda and distribution sample
simulations <- 1000
lambda <- 0.2
samples <- 40

# generate a dataset of random exponential means
mean_samples = NULL
set.seed(283764)
for (i in 1:simulations) mean_samples = c(mean_samples, mean(rexp(samples,lambda)))
```

The histogram below plots the sample distribution with the red line marking the mean.

```{r}
hist(mean_samples, main = "Sample distribution of averages")
abline(v = mean(mean_samples), lwd = 2, col = "red")
```

## Sample Mean versus Theoretical Mean

Are the mean of the distribution and the theoretical mean similar?

```{r}
mean_theo <- 1/lambda
mean_dist <- mean(mean_samples)
```

The theoretical mean of the distribution is **`r mean_theo`**, which is very close to the actual mean of the distribution, which is **`r mean_dist`**.  Given the distribution of averages is very similar to a normal distribution, this is to be expected.

## Sample Variance versus Theoretical Variance

The **theoretical variance** is defined as $Var(\bar X) = \sigma^2/n$

The **sample variance** is defined as $S^2 = \frac{\sum_{i=1}^n (X_i - \bar X)^2}{n-1}$

In this case the R function `var()` is used to calculate the variable of the sample distribution.

```{r}
var_theo <- ((1/lambda)^2)/samples
var_actual <- var(mean_samples)
```

The theoretical variance is **`r var_theo`** and the actual variance is **`r var_actual`**.  Both variances are similar, which indicates the sample distribution is similar to the theoretical distribution.

## Distribution

In the figure below you will notice the similarity between the normal distribution, in red, and the sample distribution, in blue.  In this simulation the distribution is approximately normal.

```{r}
hist(mean_samples, probability = TRUE, density=10, col="lightgray", main = "Sample distribution of average random exponentials")
abline(v = mean(mean_samples), lwd=2, col="green")
# normal distribution
curve(dnorm(x, mean = mean_theo, sd = sqrt(var_theo)), min(mean_samples)-1, max(mean_samples), col = 'red', add = TRUE)
# actual distribution 
lines(density(mean_samples), col='blue')
legend(x = 6, y = 0.5, c('Normal','Actual'), cex=0.8, col = c('red','blue'), lty = 1)
```

To compare the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials we require a new dataset of random exponentials.

```{r}
# create a similar sized dataset of random exponentials
set.seed(283764)
random_samples <- rexp(simulations, lambda)

hist(random_samples, probability = TRUE, main = "Sample distribution of random exponentials")
abline(v = mean(random_samples), lwd = 2, col = "green")

# add the sample distribution
curve(dexp(x, lambda), 0, 30, col = 'blue', add = TRUE)
```

The histogram above demonstrates the distribution of a large collection of random exponentials is not normally distributed.

Comparing the two histograms aboves demonstrates that a distribution of a large collection of averages of 40 exponentials is approximately normal and a distribution of a large collection of random exponentials is not.