---
title: "A comparison of Exponential Distribution and the Central Limit Theorem"
author: "Paul Hickman"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: false
---


## Overview

An overview of ... (2-3) sentences

Investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution is simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. For this investigation, lambda is 0.2 for all of the simulations.  The simulation will investigate the distribution of averages of 40 exponentials a thousand times.

## Simulation

_Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials._

Generate the simulation data in order to compare the sample mean and compare it to the theoretical mean of the distribution.

```{r}
# number of simulations, lambda and distribution sample
nosim <- 1000
lambda <- 0.2
sampledist <- 40

samplemean = NULL
set.seed(283764)
for (i in 1:nosim) samplemean = c(samplemean, mean(rexp(sampledist,lambda)))
```

_1. Show the sample mean and compare it to the theoretical mean of the distribution._

Plot the simulated distribution with the red line marking the mean for the distribution.

```{r}
hist(samplemean, main = "Distribution of averages")
abline(v = mean(samplemean), lwd=2, col="red")
```

### The Mean

Calculate the theoretical and actual mean of the distribution.

```{r}
mean_theo <- 1/lambda
mean_dist <- mean(samplemean)
```

The theoretical mean of the distribution is **`r mean_theo`**, which is very close to the actual mean of the simulated distribution, which is **`r mean_dist`**.

### The Variance

_2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution._

explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes

_3. Show that the distribution is approximately normal._

In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials. 

As a motivating example, compare the distribution of 1000 random uniforms
hist(runif(1000))
and the distribution of 1000 averages of 40 random uniforms

mns = NULL
for (i in 1 : 1000) mns = c(mns, mean(runif(40)))
hist(mns)
This distribution looks far more Gaussian than the original uniform distribution!


This exercise is asking you to use your knowledge of the theory given in class to relate the two distributions.  
Confused?  Try re-watching video lecture 07 for a starter on how to complete this project.




Of course, there are multiple ways one could structure a report to address the requirements above.  However, the more clearly you pose and answer each question, the easier it will be for reviewers to clearly identify and evaluate your work. 

Sample Mean versus Theoretical Mean: Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.

Sample Variance versus Theoretical Variance: Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.

Distribution: Via figures and text, explain how one can tell the distribution is approximately normal.
